Embodied Neuromorphic Benchmarks

### 2015-08-14 paper/paper.tex
Added Chris' comments v2

    Evaluating the effectiveness and performance of neuromorphic hardware is
    difficult.  It is even more difficult when the task of interest is an
    embodied task; that is, a task where the output from the neuromorphic
    hardware affects its future input through some environment.  However, embodied situations
    are one of the primary potential uses of neuromorphic hardware.  To address
    this, we present a methodology for embodied benchmarking that makes use
    of a hybrid of real physical embodiment and a type of ``minimal'' simulation.
    Minimal simulation has been shown to lead to robust real-world performance,
    while still maintaining the practical advantages of simulation, such as
    making it easy for the same benchmark to be used by many researchers.
    These benchmarks are flexible, in that they allow researchers to explicitly
    modify the benchmark to identify particular task domains where particular
    hardware excels.  To demonstrate the method, we present a novel benchmark
    where the task is to perform motor control on an arbitrary system with
    unknown external forces.

@@

To define such a benchmark, we need to be explicit about the embodiment.  If
a robot is to be controlled, we need to be explicit
about all of the details of that robot.  What motors does it have?  How are they
configured?  How strong are they?  What sensors are there?  Where are they
placed?  How accurate are they?  However, even if these questions are
answered, there is a fundamental problem in that \emph{other researchers
need access to that exact robot}.  If a benchmark is to be widely used, other
researchers developing their own neuromorphic hardware should be able to do their own 
testing on the same benchmark system.

@@

However, \emph{not} using a real physical embodiment for testing is also
problematic.  First and foremost, without an actual real-world task, why
should we have any confidence that the performance on the benchmark is
reflective of the actual usefulness of the neuromorphic hardware?  It is
widely known that simulations of robots (or other physical systems) are
often \emph{much} easier to control and better-behaved than the real thing [TODO: ref??].  The
field of robotics is filled with algorithms that work well ``in theory,'' but
fail when run on actual hardware.  We do not want a benchmark that falls into
this trap, giving high scores to hardware that does not turn out to
functional well when generalized to real situations.

@@

@@

The above considerations seem to indicate that even though using real-world
physical hardware for benchmarking is problematic, it is still better than
using simplistic simulations which may not generalize to real tasks.  However,
we believe neuromorphic benchmarking can effectively exploit an approach 
known as \emph{Minimal Simulation}
(\citealt{Jakobi97evolutionaryrobotics}).

@@

It is worth noting that a minimal simulation only has to be a
good simulation \emph{for successful behaviour}.  That is, ``if we are evolving 
corridor following behaviour, the dynamics of the simulation might
differ wildly from those of reality if the controller hits a wall or
goes round in circles, but this does
not matter, since the controllers we are interested in transferring
across the reality gap will neither
hit walls nor go round in circles.''  [ref: jacobi thesis]  If the controller is poor, we do not 
need the simulation to be at all accurate in exactly \emph{how} that poor
behaviour is manifest.  We do not need an exact detailed physics model of
the collision between a robot and a wall, or a detailed model of what happens
to a robot arm when it starts oscillating wildly due to a poor control signal.
All we need is for the simulation to be just good enough to indicate that
things have gone wrong, and thus give a low score to that controller.  This
means that, for example, in a minimal simulation of an eight-legged walking robot,
it is not necessary to have a physics simulation that correctly models what
happens when two legs collide with each other.  Rather, if legs collide
with each other, that is an indication that the walking behaviour is very poor.
As long as that result is indicated we can greatly simplify the simulation
by not including all the details necessary to deal with these sorts of physical
interactions.

\subsection{Minimal Simulation as a Benchmark}

Given the success of this approach for evolutionary robotics, we propose
using a minimal simulation as a neuromorphic benchmark.  First, we note that
one important use of a benchmark is that by knowing how well particular
hardware performs on that benchmark, you can reasonably infer how well
that hardware will perform in other situations.  For example, if an image recognition
algorithm performs well on the MNIST hand-written digit recognition
benchmark, we can use that knowledge to guess that it may also perform well on
a different recognition task.  Of course, this inference will fail if that
algorithn has been specifically over-fit to exactly that one situtation.  For
that reason, it would be useful to have a benchmark that covers a large range
of variations on the task.  If the hardware performs well across that
variability, then it is more likely to also work in whatever new situation we
want to use it in.

To achieve this, we need software simulations of the environment for the task.
These simulations
must be fast enough to run in real time (so that they can be controlled by
real neuromorphic hardware), and they must be extremely variable.  Each time
the simulation is run, different parameters will be chosed for this variability
(so one run might have a large degree of sensor noise while the next run has
none at all; one run might have more delay in the motor response and another
might have less power available).  Being successful at the benchmark means
being successful across all this variability.

The result should be a benchmark that can be run by any researcher.  The fact
that it is a simulation means that source code can be shared, and that no
specialized hardware is needed.  Furthermore, the variability in the simulation
itself can be controlled, and this can help give a rich characterization of
the benchmarked hardware.  For example, some hardware might only work with
small amounts of sensor noise, or other hardware might only work when there
is significant delay in the motor response.  This flexibility in parameters
in the benchmark allows researchers to explicitly characterize that particular
siuations where their hardware excells.

@@

The minimal simulation described above forms the core of our benchmark.
However, the point of that benchmark is that is should do a reasonable job
of generalizing to real-world physical tasks.  It is then useful to
supplement that simulation benchmark with at least one easy-to-construct 
physical analog.  This physical version would be one particular instance of the
type of situation the benchmark is meant to cover.  For that reason, it
is much more restrictive in terms of what general conclusions can be drawn
from how well different hardware performs in that situation.  Rather, it is
gives an explicit and understandable double-check that models that perform
well on the simulation benchmark also perform well in a physical environment.
Even considering the advantages of benchmarking using minimal simulations,
it is still useful to have a real physical point of comparison as well.

For this physical aspect of the benchmark, we recommend cheap, widely-available
components.  This allows a greater chance for other researchers to have
access to the same (or similar) hardware.  For the particular example 
benchmark described in the next section, we use the Lego Mindstorms EV3 kit,
a simple robotics platform available at most toy stores.

It is important to note that there is actually a theoretical advantage to
using cheap robotics hardware for benchmarking, in addition to the practical
advantages.  In particular, we \emph{don't want benchmarks that rely on
high-speed, high-accuracy devices}.  The purpose of the benchmark
is not to indicate how well this neuromorphic hardware works to control this
one particular robot in this task.  Rather, the purpose of a benchmark is to
characterise how well this neuromorphic hardware works on this task 
\emph{in general}.  The variability in the minimal simulation means that it
should be able to function across a wide variety of physical embodiments, and
so if we are to choose one particular physical embodiment to test in the
real world, then we should choose one that is not extremely high-precision.
For this reason, we believe using cheap Lego robot is actually more useful
for benchmarking than an expensive high-precision robot.\footnote{Of course,
for more complex benchmark tasks we may need sensory and motor capabiities
that are beyond that of a simple Lego robot.}

\section{A Benchmark: Adaptive Motor Control}

To demonstrate this approach to creating embodied neuromorphic benchmarks,
we now consider a basic control task.  Suppose we have a system with a number
of joints $q$ and we want to send an output $u$ to those motors such that
the joints move to a particular desired position $q_d$.  Our only output is
the signal $u$ (one for each motor) and our only inputs are the current position
of each motor $q$ and the desired positions $q_d$.

The simplest controller for such a situation is a P (proportional) controller,
where $u=K_p(q_d - q)$.  This is often supplemented with a D (derivative) term,
which helps to slow the system down as it approaches the desired position,
thus avoiding overshooting and oscillation ($u=K_p(q_d - q) + K_d(\dot{q_d} - \dot{q})$),
leading to the standard PD controller.  Both $K_p$ and $K_d$ are constants that can be
tuned to particular situations.

However, this controller has difficulty in the presence of significant external
forces.  For example, consider a single motor controlling the angle of a single
arm.  If the arm is held out to the side, gravity acting on the mass of the
arm itself will pull the arm downward.  Thus to hold the arm still at a
particular $q_d$ will require the controller to apply a force to counteract
gravity.  Since the PD controller always produces an output $u=0$ when $q=q_d$,
it cannot compensate for this, and so the arm will stay stationary at some
angle below the desired angle. [TODO: add diagram]

The standard solution to this problem is to add an I (integral) term ($K_i \int{(q_d-q) dt}$)
to the controller, making it a PID controller.  The idea here is that as the difference between where it is and
where we want it to be accumulates over time, the $K_i$ term gradually increases
how much extra force is being applied until it is large enough to counteract
the external force of gravity (or whatever other external forces are present).
However, this approach has great difficulty when $q_d$ changes, since the
external force due to gravity changes depending on the position of the arm $q$.
The controller ends up having to ``relearn'' the
correct amount of extra force needed every time $q_d$ changes.

In some robotics applications, the solution to this problem is to mathematically
analyze the geometry and mass of the system and compute exactly how much extra
force is needed.  In this particular case, the answer is straight-forward,
in that the extra torque due to gravity is $\tau=mg {l \over 2} sin(q)$, where
$m$ is the mass of the arm, $l$ is the length, and $g$ is $9.8m/s^2$.  If the
force applied by the motor is linear in $u$, then we could simply compute this
value and add it to our controller's output.  However, this assumes a perfectly
even distribution of weight in the arm, ignores momentum and other forces, and 
gets much more complex as more joints are added.  Furthermore, if this
initial computation is slightly off, or if details of the system change,
there is no way to adjust this compensation.

Fortunately, there is an adaptive solution to this problem, and it is one that
fits well with neuromorphic hardware.  \cite{Slotine1987} shows that if you
express the influence of these other external forces as $\tau=Y(q) \omega$
(where $Y(q)$ is a fixed set of functions of $q$, such as $sin(q)$, and $\omega$
is a vector of scalar weights, one for each function in $Y$), then you can learn
to compensate for these external forces by using the learning rule $\Delta \omega = \alpha Y(q) u$,
where $u$ is the basic PD control signal.

Importantly, as pointed out by \cite{SannerSlotine1992} and \cite{Lewis1996}, rather than making explicit
assumptions about the exact functions that should be in $Y(q)$, we can use
a neural network approach where each neuron is a different function of $q$.  
As long as there is enough hetereogenetity (i.e. as long as the neural activity
forms a basis space that is capable of approximating the external forces), then the
learning rule will continue to work.  This approach has been extended to
biologically plausible neurons and been used in both the
Recurrent Error-driven Adaptive Control Hierarchy model of human motor control
(\citealt{dewolf2014}) and quadcopter control (\citealt{KomerThesis}).

This then suggests an explicit neuromorphic benchmark.  The input to the
neuromorphic hardware is $q$, the system state.  This input is fed to each
neuron such that each neuron produces some output behaviour that is based on
this input.  Since $q$ will be multi-dimensional (if there is more than one
joint), we may give each neuron a random weighting of each $q$ value ($J_i=e_i \cdot q$, 
where $J_i$ is the input to neuron $i$, and $e_i$ is a randomly chosen vector\footnote{$e$ could also be chosen so as to regularly span the space of possibilities}).  Given this input, the neurons will produce some output $A$.
We now form a weighted sum of these outputs $Ad$, where $d$ is a matrix (number of neurons by number of elements in $q$)
that is initially all zeros.

To use this controller, we add its output to that of the standard PD controller.
That is, the standard controller has $u=K_p(q_d - q) + K_d(\dot{q_d} - \dot{q})$,
and so our actual output to the motor is $u + Ad$.  We then apply a learning
rule on $d$ such that $\Delta d = \alpha A \times u$.

Notice that we can think of this system as a three-layer neural
network, where the input and output layers are linear.  The first layer is $q$, the input state, one value for each joint.  The ``hidden'' layer
is $A$, the activity of a large number of neurons.  The output layer again has
one value per joint, and is the extra added signal to apply to the motors, $Ad$.
Given that this is such a canonical example of the use of neural networks, we
hope that the majority of neuromorphic hardware is flexible enough to implement
exactly this model.

@@

The one major step here that does not exist in a lot of neuromorphic hardware is
the ability to update the weights $d$.  For hardware that does have a built-in
learning rule, this rule is at least of a very common form, where the weight
update from a neuron is proportional to the activity of that neuron and an
external error signal.  This makes it an instance of the ubiquitous delta rule,
and hopefully supported by the hardware.

However, if the neuromorphic hardware being benchmarked does not have the
ability to update weights online using a learning rule of this form, then there
are two solutions.  First, the multiplication by $d$ could be done on the
output from the neuromorphic hardware.  There has to be some system to take
the neural output from the hardware and send it to the motor (or to the simulation
of the motor).  Instead of outputing the result of $Ad$, the hardware could
output $A$ (the activity of all the neurons), and the interface to the motor
could be responsible for doing the multiplication by $d$ and updating $d$
according to the learning rule.

Alternatively, we can use offline learning.  That is, rather than updating
the weights $d$ all the time, we simply record $A$ and $u$, and then after
a period of time stop the controller, compute the sum total of the changes
to $d$, load the new value of $d$ onto the neuromorphic hardware, and then
start the controller again.

@@

Now that we have defined the task, we can use the principles of Minimal
Simulation to construct a flexible and variable simulated environment for
testing adaptive control.  The idea is to make a bare-bones simulation of
the system being controlled, with built-in variability.  If the neuromorphic
controller works well across this variability, then it is likely to work well
outside of simulation as well.

The basic system variable is a vector of length $N$ holding the joint angles $q$.  
Each joint has a velocity $v$.  The force applied by each motor is 
related to the signal sent to the motor $u$, but will generally have some
maximum value $T$, so we use $tanh(u)T$. For friction, we 
simply scale the velocity by some factor $F$ every time step.  
This results in the simplistic simulation as follows:

@@

On top of this, we need to add an external perturbing force.  In a real system,
this could be the effects of gravity given the current configuration of the
motors, or of other unexpected influences.  
Rather than chosing one particular fixed external force for our benchmark,
we \emph{randomly generate} this force each time the benchmark is run.

We start with a small set of
smooth functions $f$ which are often found in dynamics equations (e.g. $x$, $x^2$,
$sin(x)$).  We then generate an external force of $K_f (\zeta \cdot f(\beta \cdot q + \gamma) + \eta)$
where $\zeta$, $\beta$, $\gamma$, and $\eta$ are all random vectors and
$K_f$ is a scaling factor to control how strong this external force is.  The
result is added to Equation 1.  Note that this means that if $q$ is 4-dimensional
(i.e. if there are four joints being controlled) and if there are three smooth
functions in $f$ (as there are here), then $\beta$, $\gamma$, and $\eta$ are 
all vectors of length 4 and $\zeta$ is a 4x12 matrix.  All of these values are
randomly chosen from the normal distribution $N(0,1)$.

Finally, we add random noise, delay, and filtering to both the input and the
output of the system.  For noise, we add $N(0,\sigma_u)$ to the control signal
$u$ and $N(0,\sigma_q)$ to the $q$ value reported back to the controller.  We
also use a low-pass filter to smooth both values (with time constants $\tau_u$
and $\tau_q$) after this noise is added.  Finally, both $q$ and $u$ are delayed
by an amount of time $t_q$ and $t_u$.

The result is not meant to be a simulation of a particular physical embodiment.
Rather, the variability in this simulation is meant to be extremely fast
to simulate, and to make it hard to ``cheat''
to control it.  In other words, if a controller manages to be able to control
the various randomly created minimal simulations of embodiment that are
generated with this approach, then we have reason to believe that it will also
be successful at controlling real embodiments.  With this simulation and
modern computers, we can easily simulate in real-time systems with dozens of joints and
highly complex interactions between them, and measure how well an adaptive
controller deals with these situations.

@@

Of course, we also want an indication that the minimal simulation defined
in the previous section is reasonably representative of the sorts of real-world
situations in which we might want to use these controllers.  Importantly, 
this physical instantiation does not have to exactly match one particular
parameter setting of the minimal simulation.  Rather, we want a physical system
that shares basic functional similarities to the minimal simulation defined
previously.

For example, we want the inputs to the system to act like $u$, in that a
positive number will increase some velocity $v$ which will in turn increase some
sensor value $q$.  We want there to be some sort of external force applied that
affects $q$, and we want that external force itself to be a function of $q$.
We want there to be communication delays and noise in the sensory and motor
system, and we want all of these effects to be somewhere within the extreme
ranges covered by the minimal simulation.  While this cannot prove that hardware
that is successful in simulation will always be successful in any similar
real-world task, it at least gives an existence proof that there is at least
one real-world task where it also performs reasonably.

For this demonstration, we define an easy-to-build example of a system that can be usefully controlled
by this adaptive method.  In particular, we use the Lego Mindstorms EV3 robot kit,
organized as shown in [TODO: add figure].  It consists of a single
motor, mounted in such a way that there is a significant amount of weight on
the arm itself (the weight of the motor itself).  Multiple motors can be
added, and other configurations can be considered and should also be suitable
for testing control, but here we just consider the basic case.

>>>>

    Evaluating the effectiveness and performance of neuromorphic hardware is
    difficult.  It is even more difficult when the task of interest is an
    embodied task; that is, a task where the output from the neuromorphic
    hardware affects some environment, which then in turn affects the
    hardware's future input.  However, embodied situations
    are one of the primary potential uses of neuromorphic hardware.  To address
    this, we present a methodology for generating embodied benchmarks that makes use
    of a hybrid of real physical embodiment and a type of ``minimal'' simulation.
    Minimal simulation has been shown to lead to robust real-world performance,
    while still maintaining the practical advantages of simulation, such as
    making it easy for the same benchmark to be used by many researchers.
    This method is flexible enough to allow researchers to explicitly
    modify the benchmarks to identify specific task domains where particular
    hardware excels.  To demonstrate the method, we present a set of novel benchmarks
    that focus on motor control for an arbitrary system with
    unknown external forces.

@@

To define an embodied benchmark, we need to be explicit about the embodiment.  If
a robot is to be controlled, we need to specify all of the details of that robot.  What motors does it have?  How are they
configured?  How strong are they?  What sensors are there?  Where are they
placed?  How accurate are they?  However, even if these questions are
answered, there is a fundamental problem in that \emph{other researchers
need access to that exact robot}.  If a benchmark is to be widely used, other
researchers developing their own neuromorphic hardware should be able to do their own 
testing on the same benchmark system.

@@

However, \emph{not} using a real physical embodiment for testing is also
problematic.  First and foremost, without an actual real-world task, why
should we have any confidence that the performance on the benchmark is
reflective of the actual usefulness of the neuromorphic hardware?  It is
widely known that simulations of robots (or other physical systems) are
often \emph{much} easier to control and better-behaved than the real thing [TODO: ref??].  The
field of robotics is filled with algorithms that work well ``in theory,'' but
fail when run on actual hardware.  We do not want a benchmark that falls into
this trap, giving high scores to hardware that does not turn out to
functional well when deployed in real situations.

@@

[TODO: maybe add a paragraph summarizing the horns of the dilemma so as to set up the
role of minimal simulation more?]

@@

The above considerations seem to indicate that even though using real-world
physical hardware for benchmarking is problematic, it is still better than
using simplistic simulations which may not generalize to real tasks.  However,
we do not think this is the case.  Instead, we believe neuromorphic benchmarking can effectively exploit an approach 
known as \emph{Minimal Simulation}
(\citealt{Jakobi97evolutionaryrobotics}).

@@

It is worth noting that a minimal simulation only has to be a
good simulation \emph{for successful behaviour}.  That is, ``if we are evolving 
corridor following behaviour, the dynamics of the simulation might
differ wildly from those of reality if the controller hits a wall or
goes round in circles, but this does
not matter, since the controllers we are interested in transferring
across the reality gap will neither
hit walls nor go round in circles.''  [ref: jacobi thesis]  If the controller is poor, we do not 
need the simulation to be at all accurate in exactly \emph{how} that poor
behaviour is manifest.  We do not need an exact detailed physics model of
the collision between a robot and a wall, or a detailed model of what happens
to a robot arm when it starts oscillating wildly due to a poor control signal.
All we need is for the simulation to be just good enough to indicate that
things have gone wrong, and thus give a low score to that controller.  This
means that, for example, in a minimal simulation of an eight-legged walking robot,
it is not necessary to have a physics simulation that correctly models what
happens when two legs collide with each other.  Rather, if legs collide
with each other, that is an indication that the walking behaviour is very poor.
As long as that result is indicated we can greatly simplify the simulation
by not including all the details necessary to model these complex physical
interactions.

\subsection{Minimal Simulation for Benchmarking}

Given the success of this approach for evolutionary robotics, we propose
using minimal simulation for neuromorphic benchmarks.  We would argue that
one important use of a benchmark is \emph{generalization}.  That is, by knowing how well particular
hardware performs on a benchmark, you can reasonably infer how well
that hardware will perform in other situations.  For example, if an image recognition
algorithm performs well on the MNIST hand-written digit recognition
benchmark, this suggests that it may also perform well on
a different recognition task.  Of course, this inference will fail if that
algorithn has been specifically over-fit to that situtation.  For
that reason, it is useful to have benchmarks that covers a wide range
of variations on the task.  If the hardware performs well across that
variability, then it is more likely to also work in sufficiently similar new situations.

To achieve this kind of transfer, we need software simulations of the environment for the task.
These simulations
must be fast enough to run in real time (so that they can be controlled by
real neuromorphic hardware), and they must be extremely variable, to encourage
robustness of the methods being benchmarked.  Each time
the simulation is run, different parameters will be chosen to give significant variability
(so one run might have a large degree of sensor noise while the next run has
none at all; one run might have more delay in the motor response and another
might have less power available).  Being successful at the benchmark means
being successful across all this variability.

The result is a benchmark that can be used by any researcher.  The fact
that it is a simulation means that source code can be shared, and that no
specialized hardware is needed.  Furthermore, the variability in the simulation
itself can be controlled, and this can help give a rich characterization of
the benchmarked hardware.  For example, some hardware might only work with
small amounts of sensor noise, while other hardware might be most effective when there
is significant delay in the motor response.  This flexibility in parameters
in the benchmark allows researchers to explicitly characterize those particular
situations where their hardware excels.

@@

The minimal simulation described above forms the core of our method for generating benchmarks.
The purpose of these benchmarks is that is that they should do a reasonable job
of generalizing to real-world physical tasks.  Consequently, it is important to
supplement simulation benchmarks with at least one easy-to-construct 
physical analog.  This physical version would be one particular instance of the
type of situation the benchmarks are meant to cover.  For that reason, the physical task
is much more restrictive in terms of what general conclusions can be drawn
from how well different hardware performs in that situation.  Rather, the purpose is to give
an explicit double-check that hardware that performs
well on the simulation benchmarks also perform well in a physical environment.

To keep the physical aspect simple, we recommend cheap, widely-available
components.  This allows a greater chance for other researchers to have
access to the same (or similar) hardware.  For the particular example 
benchmark described in the next section, we use the Lego Mindstorms EV3 kit,
a simple robotics platform available at most toy stores.

It is important to note that there is a theoretical advantage to
using cheap robotics hardware for benchmarking, in addition to the practical
advantages.  In particular, we \emph{do not want benchmarks that rely on
high-speed, high-accuracy devices}.  The purpose of benchmarks
is not to indicate how well this neuromorphic hardware works to control this
particular robot in this task.  Rather, the purpose of a benchmark is to
characterise how well some specific neuromorphic hardware works on a task 
\emph{in general}.  The variability introduced in the minimal simulation means that the hardware
should be able to function across a wide variety of physical embodiments, and
so if we are to choose one particular physical embodiment to test in the
real world, then we should choose one that is not high-precision.
For this reason, we believe using cheap Lego robots is actually more informative
for benchmarking than expensive high-precision robots.\footnote{Of course,
for more complex benchmark tasks we may need sensory and motor capabiities
that are beyond that of a simple Lego robot.}

\section{Example: Adaptive Motor Control}

To demonstrate this approach to creating embodied neuromorphic benchmarks,
we now consider a basic control task.  Suppose we have a system with a number
of joints with positions $q = [q_1, q_2, ..., q_n]$ and we want to send a control signal $u = [u_1, u_2, ..., u_n]$ to the motors at each joint
such that
the joints move to a particular desired position $q_d = [q_{d,1}, q_{d,2}, ..., q_{d,n}]$.  The only output from the controller is
the signal $u$ and the inputs are the current position
of each motor $q$ and the desired positions $q_d$.

The simplest controller for such a situation is a P (proportional) controller,
where $u=K_p(q_d - q)$.  This is often supplemented with a D (derivative) term,
which helps to slow the system down as it approaches the desired position,
thus avoiding overshoot and oscillation. This combination of terms leads to the
standard PD controller $u=K_p(q_d - q) + K_d(\dot{q_d} - \dot{q})$.
Both $K_p$ and $K_d$ are constants that can be
tuned to particular situations.

However, this controller has difficulty in the presence of significant external
forces.  For example, consider a single motor controlling the angle of a single
arm.  If the arm is held out to the side, gravity acting on the mass of the
arm itself will pull the arm downward.  Thus to hold the arm still at a
particular $q_d$ will require the controller to apply a force to counteract
gravity.  Since the PD controller always produces an output $u=0$ when $q=q_d$,
it cannot compensate for gravity, and so the arm will stay stationary at some
angle below the desired angle. [TODO: add diagram]

The standard solution to this steady-state error is to add an I (integral) term ($K_i \int{(q_d-q) dt}$)
to the controller, making it a PID controller.  As the difference between where it is ($q$) and
where we want it to be ($q_d$) accumulates over time, the $K_i$ term will gradually increase the
the extra controlled force $u$ that is being applied until it is large enough to counteract
the external force of gravity (or whatever other external forces are present).
However, this approach has great difficulty when $q_d$ changes, since the
external force due to gravity changes depending on the position of the arm $q$.
The controller ends up having to ``relearn'' the
correct amount of extra force needed every time $q_d$ changes.

In some robotics applications, this problem is solved by mathematically
analyzing the geometry and mass of the system to compute exactly how much extra
force is needed.  In this particular case, the answer is straight-forward,
in that the extra torque due to gravity is $\tau=mg {l \over 2} sin(q)$, where
$m$ is the mass of the arm, $l$ is the length, and $g$ is $9.8m/s^2$.  If the
force applied by the motor is linear in $u$, then we could simply compute this
value and add it to our controller's output.  However, this assumes a perfectly
even distribution of weight in the arm, ignores momentum, friction, and other forces, and 
gets much more complex as more joints are added.  Furthermore, if this
initial computation is slightly off, or if details of the system change,
there is no way to adjust this compensation.

Fortunately, there is an adaptive solution to this problem, and it is one that
fits well with neuromorphic hardware.  \cite{Slotine1987} show that if you
express the influence of the external forces as $\tau=Y(q) \omega$
(where $Y(q)$ is a fixed set of functions of $q$, such as $sin(q)$, and $\omega$
is a vector of scalar weights, one for each function in $Y$), then you can learn
to compensate for these external forces by using the learning rule $\Delta \omega = \alpha Y(q) u$,
where $u$ is the basic PD control signal.

Importantly, as pointed out by \cite{SannerSlotine1992} and \cite{Lewis1996}, rather than making explicit
assumptions about the exact functions that should be in $Y(q)$, we can use
a neural network approach where each neuron is a different function of $q$.  
As long as there is enough hetereogenetity (i.e. as long as the neural activity
forms a basis space that is capable of approximating the external forces), then the
learning rule will continue to work.  This approach has been extended to
biologically plausible neurons and been used in both the
Recurrent Error-driven Adaptive Control Hierarchy (REACH) model of human motor control
(\citealt{dewolf2014}) and quadcopter control (\citealt{KomerThesis}).

These considerations suggest that there is a neuromorphic-friendly family of 
algorithms to address the general problem of controlling a wide variety of 
physical systems.  Identifying those algorithms will allow us to benchmark 
their performance across example tasks and physical configurations.  
To implement these algorithms, the input to the
neuromorphic hardware is $q$, the system state.  This input is fed to each
neuron such that each neuron produces some output activity that is based on
this input.  Since $q$ will be multi-dimensional (if there is more than one
joint), we may give each neuron a random weighting of each $q$ value ($J_i=e_i \cdot q + b_i$, 
where $J_i$ is the input to neuron $i$, and $e_i$ is a randomly chosen 
vector\footnote{$e$ could also be chosen so as to regularly span the space of possibilities,
or could be learned using some back-propogation of error method.  Here, for simplicity, we only
consider the approach of randomly selecting $e_i$ and $b_i$.},
and $b_i$ is a randomly chosen bias term).  
Given this input, the neurons will produce some output $A$.
We now form a weighted sum of these outputs $Ad$, where $d$ is a matrix (number of neurons by number of elements in $q$)
that is initially all zeros.

To use this controller, we add its output to that of the standard PD controller.
That is, the standard controller has $u=K_p(q_d - q) + K_d(\dot{q_d} - \dot{q})$,
and so our actual output to the motor is $u + Ad$.  We then apply a learning
rule on $d$ such that $\Delta d = \alpha A \times u$.  Here, $\alpha$ is a learning
rate and the cross product is used so that we are applying the learning rule
on all the joints simultaneously.

Notice that we can think of this system as a three-layer neural
network, where the input and output layers are linear.  The first layer is $q$, the input state, one value for each joint.  The ``hidden'' layer
is the neurons producing activities $A$, the activity of a large number of neurons.  The output layer again has
one value per joint, and is the extra added signal to apply to the motors, $Ad$.
Given that this is a canonical example of the use of neural networks, we
expect that the majority of neuromorphic hardware is flexible enough to implement
this model.  Importantly, it functions well with spiking neuron models as 
well as non-spiking ones.  For spiking neurons, we consider $A$ to be the
instantaneous measure of the output of a neuron (i.e. whether or not it is currently outputing a spike),
filtered through a low-pass filter.  Further discussion of this sort of learning
rule and comparison to biological spiking neurons is given by [ref?].

@@

The rule for modifying the weights $d$ described here is
of a very common form, as the weight
update from a neuron is proportional to the activity of that neuron and an
external error signal.  This makes it an instance of the ubiquitous delta rule.
Thus, neuromorphic hardware that has built-in learning will often be able to natively support this rule.
However, some neuromorphic hardware does not intrinsically have
the ability to update connection weights in this manner.

In that case, there are at least two possible ways to implement this algorithm.
First, the multiplication by $d$ can be done on the
output from the neuromorphic hardware.  Any embodied neuromorphic system will have some method that takes
the neural output from the hardware and sends it to the motors (or to the simulation
of the motors).  Instead of sending the result of $Ad$, the hardware could
send $A$ (the activity of all the neurons), and the interface to the motor
can be responsible for doing the multiplication by $d$ and updating $d$
according to the learning rule.

Alternatively, it may be possible to use offline learning.  That is, rather than updating
the weights $d$ during the simulation, we record $A$ and $u$, and after
a period of time stop the controller, compute the sum of the changes
to $d$, load the new value of $d$ onto the neuromorphic hardware, and
start the controller again.

Given this variety of options for implementing adaptive algorithms of this type, 
we believe it should be possible to benchmark most neuromorphic systems on 
adaptive control tasks in this manner.

@@

Now that we have defined the task domain, we can use the principles of minimal
simulation to construct a flexible and variable simulated environment for
testing adaptive control.  In this case, we would like to develop a bare-bones simulation of
the system being controlled, with significant variability.  If the neuromorphic
controller works well across this variability, then it is likely to work well
outside of simulation as well.

The basic system variable is a vector of joint angles $q$.  
Each joint has a velocity $v$.  The force applied by each motor is 
related to the signal $u$ sent to the motors, but will generally have some
maximum value $T$, so we use $tanh(u)T$ to determine the force as a function of the control signal. To account for friction, we 
scale the velocity by some factor $F$ every time step.  
This results in the simplistic simulation described by:

@@

In addition, we add a external perturbing forces.  In a real system,
this could be the effects of gravity given the current configuration of the
motors, or of other unexpected influences.  
Rather than chosing one particular fixed external force for our benchmark,
we \emph{randomly generate} this force each time the benchmark is run.  This
ensures that the benchmark covers a wide range of possible external forces
and motor configurations, rather than just one particular situation.

Specifically, to generate this force, we start with a small set of
smooth functions $f$ which are often found in dynamics equations ($x$, $x^2$,
$sin(x)$).  We then generate an external force of $K_f (\zeta \cdot f(\beta \cdot q + \gamma) + \eta)$
where $\zeta$, $\beta$, $\gamma$, and $\eta$ are all random vectors and
$K_f$ is a scaling factor to control how strong the external force is.  The
result is added to Equation 1.  For example, if $q$ is 4-dimensional
(i.e. if there are four joints being controlled) and if there are three smooth
functions in $f$, then $\beta$, $\gamma$, and $\eta$ are 
all vectors of length 4 and $\zeta$ is a 4x12 matrix.  To introduce
significant variability, all of these values are
randomly chosen from the normal distribution $N(0,1)$.

To complete the simulation, we introduce additional sources of variability: random noise, delay, and filtering to both the input and the
output of the system.  For noise, we add $N(0,\sigma_u)$ to the control signal
$u$ and $N(0,\sigma_q)$ to the $q$ value reported back to the controller.  We
also use a low-pass filter to smooth both values (with time constants $\tau_u$
and $\tau_q$) after this noise is added, giving a damping effect.  Finally, both $q$ and $u$ are delayed
by an amount of time $t_q$ and $t_u$ to reflect communication delays that are
common in physical systems.

The resulting simulation is not meant to be an accurate portrayal of a particular physical embodiment.
Rather, this simulation is meant to be extremely fast
to simulate, and it is meant to be similarly difficult to control as a real system.
In other words, if a controller manages to be able to control
the various randomly created minimal simulations of embodiment that are
generated with this approach, then we have reason to believe that it will also
be successful at controlling real embodiments.  With this minimal simulation and
modern computers, we can run real-time simulations of systems with dozens of joints that have
highly complex interactions between them.  Consequently, we can effectively
benchmark how well an adaptive
controller deals with these situations.

@@

It is important to ensure that the minimal simulation defined
in the previous section is representative of the sorts of real-world
situations in which we want to use these same controllers.  Importantly, 
this physical instantiation does not have to exactly match any particular
parameter setting of the minimal simulation.  Rather, we want a physical system
that shares basic functional similarities to the minimal simulation defined
previously.

For example, we want the inputs to the system to act like $u$, in that a
positive number will increase some velocity $v$ which will in turn increase some
sensor value $q$.  We want there to be some sort of external applied force that
affects $q$, and we want that external force itself to be a function of $q$.
We want there to be communication delays and noise in the sensor and motor
systems, and we want all of these effects to be somewhere within the
ranges covered by the minimal simulation.  While implementing this kind of
hardware analog cannot guarantee that neuromorphic hardware
that is successful in simulation will be successful in every similar
real-world task, it does provide an existence proof that there is at least
one real-world task where the hardware performs similarly to how it performs in simulation.

For our specific demonstration, we describe an easy-to-build system that can be usefully controlled
by this adaptive method.  In particular, we use the Lego Mindstorms EV3 robot kit,
organized as shown in [TODO: add figure].  It consists of a single
motor, mounted in such a way that there is a significant amount of weight on
the arm itself (i.e., the weight of the motor itself).  Multiple motors can be
added, and other configurations can be considered and should also be suitable
for testing control, but here we consider only this basic case.
